---
title: "Assignment 6"
author: "Mia Kuntz"
date: "20 December 2022"
output: html_document
---

# Installing packages
I first have to install the necessary packages and libraries for my code to work.
```{r libraries}
library(tidyverse)
library(here)
library(lubridate)
```


# 1. Task - Calculating GDP
## Loading the data set and checking class
In order to complete the first task of the assignment, I first have to read the data *gaminder_data* into my script. I check the working directory before reading in the csv file, and thereafter check the class of the file when loaded. 
```{r loading data}
getwd()
gapminder <- read.csv("gapminder_data.csv")
class(gapminder)
```
The class shows as a data frame.

## Defining GDP function using defensive functions
The following code is the defensive function I will use to calculate GDP for different nations.
```{r defining function}
calcGDP <- function(dat, year=NULL, country=NULL) {
  if(!is.null(year)) {
    dat <- dat[dat$year %in% year, ]
  }
  if (!is.null(country)) {
    dat <- dat[dat$country %in% country,]
  }
  gdp <- dat$pop * dat$gdpPercap

  new <- cbind(dat, gdp=gdp)
  return(new)
}
```

## Calculating the GDP of Denmark in the years 1967, 1977, 1987, 1997, 2007, and 2017
By using the defensive function above, I am able to calculate the GDP of Denmark in the given years.
```{r DK GDP}
calcGDP(gapminder, year = c(1952, 1977, 1987, 1997, 2007), country = "Denmark")
```
It is worth noting that the year 2017 was given as a task, but could not be calculated due to data not being available.

# 2. Task - Creating script
## Assigning objects
In order to create the script given in the 2. task of the assignment, I first have to create different objects pertaining to the unique identifications we need in the script. 
```{r assigning objects}
candidateCountries <- grep("^B", unique(gapminder$country), value=TRUE)
lowerThreshold <- 50
upperThreshold <- 70
```

## Loop
Following script uses loops to check for the given conditions of the task. 
```{r loop}
for (iCountry in candidateCountries){
  tmp <- mean(gapminder[gapminder$country == iCountry, "lifeExp"])

  if(tmp < lowerThreshold){
    cat("Average Life Expectancy in", iCountry, "is less than", lowerThreshold, "\n")
  }
  else if(tmp > lowerThreshold && tmp < upperThreshold){
    cat("Average Life Expectancy in", iCountry, "is between", lowerThreshold, "and", upperThreshold, "\n")
  }
  else{
    cat("Average Life Expectancy in", iCountry, "is greater than", upperThreshold, "\n")
  }
  rm(tmp)
}
```
This shows whether  countries which starts with B in the *gapminder_data* either has a life expectancy of under 50, between 50 and 70, or above 70. 




